---
title: "state_profile"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this assignment.

```{r}
library(tidycensus)
library(tidyverse)
library(janitor)
library(sf)
library(tigris)
library(rvest)
library(lubridate)
```

## Initial Exploration

**Q1.**. Write R code that generates some basic descriptive statistics that allows you to describe the applications from your state. This code should produce results that describe the data in different ways: how many applications there are, along with ways to understand the typical and most frequent values for columns you find interesting or newsworthy. You should produce at least five (5) descriptive statistics and write up a summary of the data using them.

**A1.** Ohio had 352,010 PPP loan applications. I started by creating a separate dataframe that was a bit cleaner. That dataframe had standardized, five-digit zip codes and city and address names. With that dataframe, I started doing my analysis, focusing on specific categories. 

I started on a larger scale looking at Ohio cities. This wasn't a particularly interesting finding because the top three were predictably Columbus, Cleveland and Cincinnati. Still, it confirmed  what I thought to be true, so that was good to know.

I then wanted to go deeper and look at counties. I ran into similar results tha I had with the cities dataframe; the top counties had the largest cities in them. The counties were CUYAHOGA (Cleveland), FRANKLIN (Columbus) and HAMILTON (Cincy). Where things did get a little interesting was when we added an average PPP amount column. When we did that, HOLMES (111924.81) and TUSCARAWAS (105982.16) County jumped to the front. Holmes was 61 of 88 counties in total loans and Tuscarawas was 32. The population of Holmes is about 44,000 and Tuscarawas is 92,000. The two largest industries in Tuscarawas are Manufacturing and Health Care & Social Assistance, according to https://datausa.io/profile/geo/tuscarawas-county-oh. Both are industries that were impacted by the pandemic, so it's possible they applied for and got large loans. Holmes' top two industries were Manufacturing and Retail Trade. The explanation here in PPP loans is similar to that of Tuscarawas County. 

Another interesting data point to note is that the industry with the most loans was beauty salons (16,955). With people not going in to get beauty work, it makes sense why this industry would be applying for PPP loans. 

While a vast majority of the businesses in the state were non-franchise operations, there were some chains that received loans. Of the franchise businesses, Subway locations got 396 loans. McDonald's got 172 and General Motors got 85. It's interesting that two franchises were restaurants and one was car manufacturer. While different industries, both were slammed by the pandemic.

Another finding developed was which zip codes had the most loans. While that result in and of itself is decently interesting, we can make it more interesting by simply googling the median incomes for those zip codes. The question here is did poorer areas get more loans? Or did the program favor wealthier areas? This is a good way to do it because it would be silly to just look at which zip code got the most loans. If we did that, the biggest cities would be on top and we wouldn't learn a whole lot. We also have to remember that someone could have put a zip code for their businesses as opposed to their home address, so these numbers won't be exact. Still, it does give some insight into what areas were getting loans and if there was any bias from lenders based on income. Here are the top five zip codes with median income: 44128 (32,691), 44122 (74,054), 44105 (25,441), 43229 (37,518), 44120 (29,643). Four are relatively poor areas, while one is middle class. This suggests that the PPP was relatively successful because people in lower income brackets were disproportionately impacted by the pandemic. Therefore, they needed more assistance from the PPP.

The last bit I wanted to look at was towns where undergraduate institutions were located, specifically where I went and my brother goes, so I picked Licking County (Denison University) and Lorain County (Oberlin College). I then looked at the the individual towns - Granville for Denison and Oberlin for Oberlin. Interestingly, neither college had a single PPP loan, according to the data. With colleges losing large amounts of money because of the pandemic - nobody was paying room and board - I imagined they would be applying for PPP loans. One potential explanation for this is that both colleges have endowments over 900 million dollars, so maybe they used that to help during the pandemic. Another is that perhaps the SBA was weary of giving loans, many of which were forgiven, to large institutions with copious amounts of funds for a program that was designed for small businesses, which colleges certainly are not.

```{r}
ohio_ppp <- read_csv('/Users/Ian/Downloads/ohio_ppp.csv')

ohio_ppp_5_zip <- ohio_ppp %>%
  mutate(zip = str_sub(zip, 1, 5)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(address = str_to_title(address))

ohio_cities <- ohio_ppp_5_zip %>%
  group_by(city) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

ohio_counties <- ohio_ppp_5_zip %>%
  group_by(project_county_name, state) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount)) %>%
  arrange(desc(count)) %>%
  filter(count > 19)

naics <- ohio_ppp_5_zip %>%
  group_by(naics_code) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

ohio_zip <- ohio_ppp_5_zip %>%
  group_by(zip, state) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount)) %>%
  arrange(desc(count))

ohio_address <- ohio_ppp_5_zip %>%
  group_by(address) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount), total_jobs_retained = sum(jobs_retained)) %>%
  arrange(desc(count))

ppp_lenders <- ohio_ppp_5_zip %>%
  group_by(lender) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount)) %>%
  arrange(desc(count))

ohio_franchise <- ohio_ppp_5_zip %>%
  group_by(franchise_name) %>%
  summarise(count=n(), total = sum(amount)) %>%
  arrange(desc(count))

na_naics <- ohio_ppp_5_zip %>%
  filter(is.na(naics_code)) %>%
  arrange(desc(amount))

#I was just interested in Cuyahoga. There isn't anything particularly revelatory here.
cuyahoga <- ohio_ppp_5_zip %>%
  filter(project_county_name == "CUYAHOGA") %>%
  mutate(city = str_to_title(city)) %>%
  group_by (city) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

#This is the county where Denison University (my undergrad) is located!
licking <- ohio_ppp_5_zip %>%
  filter(project_county_name == "LICKING") %>%
  mutate(city = str_to_title(city)) %>%
  group_by (city) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

granville <- ohio_ppp_5_zip %>%
  mutate(city = str_to_title(city)) %>%
  filter(city == "Granville")

granville_naics <- granville %>%
  group_by (naics_code) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
  
#This is where Oberlin College (where my brother goes) is located.
lorain <- ohio_ppp_5_zip %>%
  filter(project_county_name == "LORAIN") %>%
  mutate(city = str_to_title(city)) %>%
  group_by (city) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

oberlin <- ohio_ppp_5_zip %>%
  mutate(city = str_to_title(city)) %>%
  filter(city == "Oberlin")

oberlin_naics <- oberlin %>%
  group_by (naics_code) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. 

Then, make a county map using ggplot showing the per-capita data. 

Make a zip code map showing the difference from the statewide per-capita figure. 

Describe the most interesting or newsworthy findings based on your exploration.

**A2.** While the zip code map is a little bit jumbled, it does highlight one important piece of information that is also pertinent to the county map; the northern part of Ohio seemed to receive more loans than the central or southern parts of the state. 

The two maps I made showed the per-capita data in terms of loans based on county and zip code. The counties with the highest loans per 10,000 people were CUYAHOGA, FRANKLIN and HAMILTON. Another interesting note is that counties adjacent to CUYAHOGA, FRANKLIN and HAMILTON seemed to benefit from their neighbors' pull. We see little pockets of higher loans in counties around the top three. The closer you are to a city or large metropolitan area, the more loans per 10K you were to get Overall, the northern part of Ohio received more loans than the rest of the state, primarily the eastern part of the state. In other words, areas closer to Indiana and Minnesota received more loans than areas bordering Pennsylvania and West Virginia. The one county in the east that is the exception to lower loans is Washington County (in orange). According to https://www.seohioport.com/why-the-mov/workforce/, Washington's biggest industry is Manufacturing, which could help explain why more loans went there. Another point on Washington: the largest employer is Memorial Health System, which employs over 3,000 people, so the PPP loans could have been going to keep people employed and cover operation costs, which would have upped the amount of loans to that county.

```{r}
census_api_key("1cc6052b8d53c607d6876a2e5a3c44b607a324d4", overwrite=TRUE)

ohio_county_population <- get_acs(geography = "county", 
              variables = c(population = "B01001_001"), 
              state = "OH",
              geometry = TRUE)

ohio_population <-  get_acs(geography = "state", 
              variables = "B01001_001",
              year = 2019) %>%
              filter(NAME == "Ohio")

ohio_zip_population <- get_acs(geography = "zcta", 
              variables = c(population = "B01001_001"), 
              state = "OH",
       geometry = TRUE)

#Code for county map
ohio_county_population <- ohio_county_population %>%
  mutate(county = str_replace(NAME, " County, Ohio", "")) %>%
  mutate(county = str_to_upper(county))

ohio_ppp_county_with_ohio_pop <- ohio_counties %>%
  left_join(ohio_county_population, by=c("project_county_name" = "county")) %>%
  rename(population = estimate) %>%
  mutate(loans_per_10k = total/population*10000) %>%
  arrange(desc(loans_per_10k))

ggplot() + 
  geom_sf(data=ohio_ppp_county_with_ohio_pop, aes(fill=loans_per_10k, geometry=geometry)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") 

#Code for zip code map
ohio_ppp_zip_with_ohio_pop <- ohio_zip_population %>%
  left_join(ohio_zip, by=c("GEOID" = "zip")) %>%
  rename(population = estimate) %>%
  mutate(loans_per_10k = total/population*10000) %>%
  arrange(desc(loans_per_10k))

ggplot() + 
  geom_sf(data=ohio_ppp_zip_with_ohio_pop, aes(fill=loans_per_10k)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log")
```

## Lender Analysis

**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). 

Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. 

Describe the most noticeable outlier lenders on that graphic below.

**A3.** 
Because my dataframe did not have any undisbursed amounts, I used min and max values to get the data. 

After creating the dataframe, we see that there seem to be several outliers, primarily the lender Harvest Small Business Finance, LLC. This lender gave out 6672 loans for a total of $138,997,776. 

Our scatterplot tells us several interesting bits of information. First, it confirms that the most people applied for the max amount - pretty straightforward. For sole proprietors or independent contractors with no employees, the maximum possible PPP loan is 20,833 dollars. When we look at our chart, a majority of the loans are either that amount or close to it. 

Another thing to point out is that most of the rest of the applications fall on the right side of the graph, which means that there were more lenders who gave smaller amounts than larger amounts. One potential explanation for this could be that those lenders started loaning later in the program. The later an entity started loaning, the less total loans it would have. I know that we had spoken about how to do that, but I don't know how to write that code. 

We also see that the highest loan was 50,000 dollars. This is well below the maximum of 10 million that businesses were capped at. It's also interesting to look at the seven largest lenders. There were seven lenders - The Huntington National Bank, PNC Bank, National Association, KeyBank National Association, Fifth Third Bank, JPMorgan Chase Bank, National Association, U.S. Bank, National Association, U.S. Bank, National Association - that gave hundreds of thousands of loans worth the max (20833) over the course of this project. Notice that these banks all have large-scale operations, which makes sense as to why they gave out so much.

```{r}
ppp_lenders <- ohio_ppp_5_zip %>%
  group_by(lender, amount) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount), total_undisbursed = sum(undisbursed_amount), min = min(),   max = max(approval)) %>%
  filter(count > 100) %>%
  arrange(desc(count))

ppp_lenders %>% 
ggplot(aes(x=amount, y=count)) + 
  geom_point() +
labs(
 title="Ohio Loans by Lender",
    x = "Loan Amount",
    y = "Total Loans",)
```

## Industry Analysis

**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. To do this, you will need to add FIPS codes to your PPP dataset.

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

**A4.** Believe it or not, I really did try to figure out the last portion to this question. While I don't have a choropleth map for Ohio counties, I can try and make an educated guess as to what it would have looked like. Similarly to a lot of the other maps made for this project, I have a feeling that the hot spots would have been in the cities - Cleveland, Columbus and Cincinnati. The naics code chosen was for limited-service restaurants, and I have to imagine there are quite a few of those in the major metropolitan areas. 

```{r}
all_fips <- fips_codes %>% as_tibble()
ohio_fips <- all_fips %>% filter(state == 'OH')

ohio_fips <- ohio_fips %>% mutate(fips_code = str_c(state_code, county_code)) %>%
  mutate(county = str_replace(county, " County", "")) %>%
  mutate(county = str_to_upper(county))

ohio_counties_with_fips <- ohio_counties %>%
  left_join(ohio_fips, by=c('project_county_name'='county'))

top_10_industries <- ohio_ppp_5_zip %>%
  group_by(naics_code) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(10)

all_cbp <- read_csv('/Users/Ian/Downloads/cbp19co.txt')
ohio_cbp <- all_cbp %>% filter(fipstate == '39')

ohio_cbp <- ohio_cbp %>%
  filter(!str_detect(naics, "/")) %>%
  filter(!str_detect(naics, "-")) %>%
  mutate(fips = str_c(fipstate, fipscty))

top_10_cbp <- ohio_cbp %>%
  group_by(naics) %>%
  summarise(total_businesses = sum(est)) %>%
  arrange(desc(total_businesses)) %>%
  top_n(10)

top_10_industries_limited <- ohio_ppp_5_zip %>%
  filter(business_type != "Sole Proprietorship") %>%
  filter(business_type != "Self-Employed Individuals") %>%
  group_by(naics_code) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(10)
  
#cbp number of businesses the census bureau believes exist
#top_10_limited is the number of PPP loans per industry

#I did not know how to add fips_code to this dataframe so I could join it with naics_ohio_plus_fips.
ohio_primary_naics_code <- ohio_ppp_5_zip %>% 
  filter(naics_code == "722513") %>% 
  group_by(project_county_name, fips_code) %>% 
  summarise(total=n()) %>% 
  arrange(desc(total))

naics_ohio_plus_fips <- ohio_cbp %>%
  filter(naics == "722513") %>% 
  group_by(fips) %>% 
  summarise(total_businesses = sum(est)) %>%
  arrange(desc(total_businesses))

ohio_counties_plus_fips <- ohio_primary_naics_code %>%
  left_join(naics_ohio_plus_fips, by=c("fips"))

#dataframe with county and fips code in both cases, county name 
#in ohio_ppp_5_zip and ohio_cbp fips code
  
write_csv(ohio_counties_plus_fips, "data/ohio_counties_plus_fips.csv")

#I'm sorry I didn't generate a choropleth map. I have never felt more outside my comfort zone in a class or project. I wouldn't have completed a single step of this project without the help of others. The amount of hand-holding I needed throughout this class is embarrassing. I'm sorry I couldn't be a better student for you. You are a kind teacher who has been unbelievably supportive this semester. Again, I apologize that I couldn't give you a better final project. Have a wonderful summer!
```

## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.** The most promising story idea that I found during this exploration is that of multiple loans going to the same address. This is a great way to look at fraud, of which there was a lot over the course of this project. There were several Ohio businesses that got A LOT of loans, so that would be the starting point for a potential story.

In terms of the reporting, I would want to go to the physical addresses to look at which businesses located were located at that spot. It could be that one address houses myriad different corporations, which would explain why one address got a lot of loans.

If it was just one address, then I would start asking questions. It would be useful to talk to a manager or owner (if they would give you time), but former employees would also be good. It's possible that the company fired people even though it was getting loans. Talking to people who were impacted would be a great place to start when trying to get an understanding of the money that went to this one business. 

In terms of the actual data, one address stood out. 4503 Marburg Ave received 57 loans for $3,899,100.10. The business at this address is a Planet Fitness that, according to google, is temporarily closed. It's possible that this one gym applied for over 3 million dollars in loans, but 57 separate applications is still a lot of loans. This address is in a shopping area, so it's possible that there were other businesses that used this address. Still, why would it be temporarily closed if it received millions in PPP loans? 