---
title: "state_profile"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this assignment.

```{r}
library(tidycensus)
library(tidyverse)
library(janitor)
library(sf)
library(tigris)
library(rvest)
library(lubridate)
```

## Initial Exploration

**Q1.**. Write R code that generates some basic descriptive statistics that allows you to describe the applications from your state. This code should produce results that describe the data in different ways: how many applications there are, along with ways to understand the typical and most frequent values for columns you find interesting or newsworthy. You should produce at least five (5) descriptive statistics and write up a summary of the data using them.

**A1.** Ohio had 352,010 PPP loan applications. I started by creating a separate dataframe that was a bit cleaner. That dataframe had standardized, five-digit zip codes and city and address names. With that dataframe, I started doing my analysis.I then started looking at specific categories. I started on a larger scale looking at Ohio cities. This wasn't a particularly interesting finding becasue the top three were predictably Columbus, Cleveland and Cincinnati. I then wanted to go deeper and look at counties. I ran into a similar issue with the cities dataframe; the top counties had the largest cities in them. The countie were CUYAHOGA (Cleveland), FRANKLIN (Columbus) and HAMILTON (Cincy). Where things did get a litte interesting was when we added an avergae PPP amount column. When we did that, HOLMES (111924.81) and TUSCARAWAS (105982.16) jumped to the front. Holmes was 61 of 88 counties in total laons and Tuscarawas was 32. The population of Holmes is about 44,000 and Tuscarawas is 92,000. The two largest industries in Tuscarawas are Manufacturing and Health Care & Social Assistance, according to https://datausa.io/profile/geo/tuscarawas-county-oh. Both are industries that were imapcted by the pandemic, so it's possible these industries applied for and got large loans. Holmes' top two industries were Manufacturing and Retail Trade. The explanation here in PPP loans is similar to that of Tuscarawas County. Another interesting data point to note is that the industry with the most loans was beauty salons (16,955). With people not going in to get beauty work, it makes sense why this industry would be applying for PPP loans. I also looked at the total loans and total amounts specific addresses received. This address - 4503 Marburg Ave - received 57 loans for $3,899,100.10. The business at this address is a Planet Fitness that, according to google, is temporarily closed. It's possible that this one gyme applied for over 3 million dollars in loans, but 57 separate applications is still a lot of loans. Thsi address is in a shopping area, so it's possible that there were other businesses that used this address. While a vast majority of the businesses in the state were non-franchise operations, there were some chains that received loans. Of the franchise businesses, Subway locations got 396 loans. McDonalds got 172 and General Motors got 85. The last bit I wanted to look at was towns where undergraduate institutions were located. For this, I picked Licking County (Denison University) and Lorain County (Oberlin College). I then looked at the the individual towns - Granville for Denisn and Lorain for Oberlin. Interestingly, neither college had a single PPP loan, according to the data. With colleges losing large amount of money because of the pandemic, nobody was paying room and board, I imagined they would be applying for PPP loans. One potential explanation for this is that both colleges have endowmnets over 900 million dollars, so maybe they used that to help during the pandemic. 


```{r}
ohio_ppp <- read_csv('/Users/Ian/Downloads/ohio_ppp.csv')

ohio_ppp_5_zip <- ohio_ppp %>%
  mutate(zip = str_sub(zip, 1, 5)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(address = str_to_title(address))

ohio_cities <- ohio_ppp_5_zip %>%
  group_by(city) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

ohio_counties <- ohio_ppp_5_zip %>%
  group_by(project_county_name, state) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount)) %>%
  arrange(desc(count)) %>%
  filter(count > 19)

naics <- ohio_ppp_5_zip %>%
  group_by(naics_code) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

ohio_zip <- ohio_ppp_5_zip %>%
  group_by(zip, state) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount)) %>%
  arrange(desc(count))

ohio_address <- ohio_ppp_5_zip %>%
  group_by(address) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount), total_jobs_retained = sum(jobs_retained)) %>%
  arrange(desc(count))

ppp_lenders <- ohio_ppp_5_zip %>%
  group_by(lender) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount)) %>%
  arrange(desc(count))

ohio_franchise <- ohio_ppp_5_zip %>%
  group_by(franchise_name) %>%
  summarise(count=n(), total = sum(amount)) %>%
  arrange(desc(count))

na_naics <- ohio_ppp_5_zip %>%
  filter(is.na(naics_code)) %>%
  arrange(desc(amount))

#I was just interested in Cuyahoga. There isn't anything particularly revelatory here.
cuyahoga <- ohio_ppp_5_zip %>%
  filter(project_county_name == "CUYAHOGA") %>%
  mutate(city = str_to_title(city)) %>%
  group_by (city) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

#This is the county where Denison University (my undergrad) is located!
licking <- ohio_ppp_5_zip %>%
  filter(project_county_name == "LICKING") %>%
  mutate(city = str_to_title(city)) %>%
  group_by (city) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

granville <- ohio_ppp_5_zip %>%
  mutate(city = str_to_title(city)) %>%
  filter(city == "Granville")

granville_naics <- granville %>%
  group_by (naics_code) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
  
#This is where Oberlin College (where my brother goes) is located.
lorain <- ohio_ppp_5_zip %>%
  filter(project_county_name == "LORAIN") %>%
  mutate(city = str_to_title(city)) %>%
  group_by (city) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

oberlin <- ohio_ppp_5_zip %>%
  mutate(city = str_to_title(city)) %>%
  filter(city == "Oberlin")

oberlin_naics <- oberlin %>%
  group_by (naics_code) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. 

Then, make a county map using ggplot showing the per-capita data. 

Make a zip code map showing the difference from the statewide per-capita figure. 

Describe the most interesting or newsworthy findings based on your exploration.

**A2.** While the zip code map is a little bit jumbled, it does highlight one important piece of information that is also pertinent to the county map; the northern part of Ohio seemed to receive more loans than the central or southern parts of the state. The two maps I made showed the per-capita data in terms of loans based on county and zip code. Unsurprisingly, the counties with the highest loans per 10,000 people were...you guessed it: CUYAHOGA, FRANKLIN and HAMILTON. Another interesting note is that counties adjacent to CUYAHOGA, FRANKLIN and HAMILTON seemed to benefit from their neighbors' pull. We see little pockets of higher loans in counties around the top three. Overall, the norhtern part of Ohio received more loans than the rest of the state, primarily the eastern part of the state. In other words, areas closer to Indiana and Minnesota received more loans than areas bordering Pennsylvania and West Virginia. The one county in the east that is the exception to lower loans is Washington County (in orange). According to https://www.seohioport.com/why-the-mov/workforce/, Washington's biggest industry is Manufacturing, which explains why more loans went there. Another point on Washington: the largest employer is Memorial Health System, which employs over 3,000 people, so the PPP loans could have been going to keep people employed and cover operation costs. 

```{r}
census_api_key("1cc6052b8d53c607d6876a2e5a3c44b607a324d4", overwrite=TRUE)

ohio_county_population <- get_acs(geography = "county", 
              variables = c(population = "B01001_001"), 
              state = "OH",
              geometry = TRUE)

ohio_population <-  get_acs(geography = "state", 
              variables = "B01001_001",
              year = 2019) %>%
              filter(NAME == "Ohio")

ohio_zip_population <- get_acs(geography = "zcta", 
              variables = c(population = "B01001_001"), 
              state = "OH",
       geometry = TRUE)

#Code for county map
ohio_county_population <- ohio_county_population %>%
  mutate(county = str_replace(NAME, " County, Ohio", "")) %>%
  mutate(county = str_to_upper(county))

ohio_ppp_county_with_ohio_pop <- ohio_counties %>%
  left_join(ohio_county_population, by=c("project_county_name" = "county")) %>%
  rename(population = estimate) %>%
  mutate(loans_per_10k = total/population*10000) %>%
  arrange(desc(loans_per_10k))

ggplot() + 
  geom_sf(data=ohio_ppp_county_with_ohio_pop, aes(fill=loans_per_10k, geometry=geometry)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") 

#Code for zip code map
ohio_ppp_zip_with_ohio_pop <- ohio_zip_population %>%
  left_join(ohio_zip, by=c("GEOID" = "zip")) %>%
  rename(population = estimate) %>%
  mutate(loans_per_10k = total/population*10000) %>%
  arrange(desc(loans_per_10k))

ggplot() + 
  geom_sf(data=ohio_ppp_zip_with_ohio_pop, aes(fill=loans_per_10k)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log")
```

## Lender Analysis

**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). 

Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. 

Describe the most noticeable outlier lenders on that graphic below.

**A3.** 
Potential theory, do the lenders who stand, if they were lending throughout the life of the program

```{r}
ppp_lenders <- ohio_ppp_5_zip %>%
  group_by(lender, amount) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount), total_undisbursed = sum(undisbursed_amount), min = min(), max = max(approval)) %>%
  filter(count > 100) %>%
  arrange(desc(count))

ppp_lenders %>% 
ggplot(aes(x=amount, y=count)) + 
  geom_point() +
labs(
 title="Ohio Loans by Lender",
    x = "Loan Amount",
    y = "Total Loans",)
```

## Industry Analysis

**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. To do this, you will need to add FIPS codes to your PPP dataset.

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

**A4.** 

PRE LAB TEN
```{r}
all_fips <- fips_codes %>% as_tibble()
ohio_fips <- all_fips %>% filter(state == 'OH')

ohio_fips <- ohio_fips %>% mutate(fips_code = str_c(state_code, county_code)) %>%
  mutate(county = str_replace(county, " County", "")) %>%
  mutate(county = str_to_upper(county))

ohio_counties_with_fips <- ohio_counties %>%
  left_join(ohio_fips, by=c('project_county_name'='county'))

top_10_industries <- ohio_ppp_5_zip %>%
  group_by(naics_code) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(10)

all_cbp <- read_csv('/Users/Ian/Downloads/cbp19co.txt')
ohio_cbp <- all_cbp %>% filter(fipstate == '39')

ohio_cbp <- ohio_cbp %>%
  filter(!str_detect(naics, "/")) %>%
  filter(!str_detect(naics, "-"))

top_10_cbp <- ohio_cbp %>%
  group_by(naics) %>%
  summarise(total_businesses = sum(est)) %>%
  arrange(desc(total_businesses)) %>%
  top_n(10)

top_10_industries_limited <- ohio_ppp_5_zip %>%
  filter(business_type != "Sole Proprietorship") %>%
  filter(business_type != "Self-Employed Individuals") %>%
  group_by(naics_code) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(10)
  
#cbp number of businesses the census bureau believes exist
#top_10_limited is the number of PPP loans per industry


primary_naics <- ohio_ppp_5_zip %>%
  filter(naics_code == "722513") %>%
  group_by(project_county_name, state, naics_code) %>%
  summarise(count = n(), total = sum(amount), average = mean(amount)) %>%
  arrange(desc(count))

primary_naics_in_cbp <- ohio_cbp %>%
  filter(naics_code == "722513") %>%
  group_by(naics) %>%
  summarise() %>%
  arrange(desc())

#dataframe with county and fips code in both cases, county name 
  #in ohio_ppp_5_zip and ohio_cbp fips code
  
  

write_csv(ohio_counties_plus_fips, "data/ohio_counties_plus_fips.csv")

```

## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.**
