---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into the following options - under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus - and calculates how many loans are in each group. How would you describe the distribution of loans in a sentence?
Hint: You will need to use `case_when` to make this work.
**A1** [write your answer here] The most loans are in the 10000_to_50000 range (15089). The lowest group was laons above 1 million (356).

```{r}
# Write the code to produce your answer here
amount_groups <- west_virginia_ppp %>%
mutate(
    amount = case_when(
    amount < 10000 ~ 'under_10000',
    amount >= 10000 & amount < 50000 ~ '10000_to_50000',
    amount >= 50000 & amount < 100000 ~ '50000_to_100000',
    amount >= 100000 & amount < 500000 ~ '100000_to_500000',
    amount >= 500000 & amount < 1000000 ~ '500000_to_1000000',
    amount >= 1000000 ~ '1000000'))%>%
group_by(amount) %>%
  summarize(count = n())

```

**Q2** Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** [write your answer here] The servicing lender with the most loans with undisbursed amounts greater than 0 is Davis Trust Company. Miners & Merchants Bank also has one undisbursed amounts greater than 0.

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>% filter(undisbursed_amount > 0) %>%
 group_by(servicing_lender_name) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

**Q3** What are some potential reasons why a loan might not be paid out? What questions does that circumstance raise for you? 
**A3** [write your answer here] 1. The main issue I can think of is that the lender ran out of funds/did not have enough to give all of the applicants. The primary question this raises is why certain lenders agree to be a part of the program. In other words, why would Davis Trust Company agree to give out loans if it did not have the funds?


**Q4** Which month in 2021 had the most loans approved?
**A4** [write your answer here] March 2021 had the most loans approved with 4394.

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) %>%
  filter(Month > "2020-12-01")
#for filters(column (Month) comparison (>) value ("2020-12-01"))
```

**Q5** How many loans went to West Virginia applicants that listed the maximum number of jobs retained (500)? Which one had the earliest approved date?
**A5** [write your answer here]

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>% 
  filter(jobs_retained == 500) %>%
  arrange(date_approved)
```
