---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned 01/27

The command to set a working directory is setwd(). For example:

```{r}
   setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
  install.packages("tidyverse")
  library(tidyverse)
```

#how to filter for when a certain result is "NA":
filter(is.na(forgiveness_date)) %>%

# we changed the data type of the "id" field from a number (dbl) to a character (chr).
ppp_maryland_loans <- ppp_maryland_loans %>%
  mutate(id = as.character(id))

glimpse(ppp_maryland_loans)

#How to look for date gaps in data
ppp_maryland_loans %>%
  mutate(month_year_approved = floor_date(date_approved, "month")) %>%
  group_by(month_year_approved) %>%
   summarise(
    count=n()
  ) %>%
  arrange(month_year_approved)

