---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned 01/27

The command to set a working directory is setwd(). For example:

```{r}
   setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
  install.packages("tidyverse")
  library(tidyverse)
```

#how to filter for when a certain result is "NA":
filter(is.na(forgiveness_date)) %>%

# we changed the data type of the "id" field from a number (dbl) to a character (chr).
ppp_maryland_loans <- ppp_maryland_loans %>%
  mutate(id = as.character(id))

glimpse(ppp_maryland_loans)

#How to look for date gaps in data
ppp_maryland_loans %>%
  mutate(month_year_approved = floor_date(date_approved, "month")) %>%
  group_by(month_year_approved) %>%
   summarise(
    count=n()
  ) %>%
  arrange(month_year_approved)

##For cleaning datasets with inconsistencies
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names()
# display the cleaned dataset
cleaned_arnold_md_loans

##Change "x1_id" to "id"
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id)
# display the cleaned dataset
cleaned_arnold_md_loans

##For removing duplicates
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct()
# display the cleaned dataset
cleaned_arnold_md_loans

##For standardizing zip codes
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))
# display the cleaned dataset
cleaned_arnold_md_loans

##To clean up spelling
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city))
# display the cleaned dataset
cleaned_arnold_md_loans

##Clean up city field more with case_when()
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    city == "Anold" ~ "Arnold",
    str_detect(city,"^Arnl") ~ "Arnold",
    TRUE ~ city
  ))
# display the cleaned dataset
cleaned_arnold_md_loans

#Using bind rows

# bind_rows with list
```{r}
estimates <- bind_rows(list(popestimate_2010, popestimate_2015, popestimate_2020))
View(estimates)
```

#Assign Joined Data to New Dataframe

```{r}
maryland_ppp_with_naics <- maryland_ppp %>% left_join(naics_codes, by="naics_code")
```

Fix ZCTA Data

```{r}
maryland_zcta <- maryland_zcta %>% mutate(across(ZCTA5N, as.character))
```

#Basic Syntax and Order
my_result <- original_data %>%
  filter(some_column == "some value") %>%
  mutate(some_column=str_to_title(some_column)) %>%
  group_by(some_column) %>%
  summarise(new_aggregate = n()) %>%
  arrange(desc(new_aggregate))
  
