---
title: "lab_11"
author: "sean mussenden"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. You'll need tidyverse, tidycensus, janitor, sf and tigris, and you'll also need to install your census API key.

```{r}
library(tidyverse)
library(tidycensus)
library(janitor)
library(sf)
library(tigris)
```

**Task** set your API Key

```{r echo=FALSE}
census_api_key("1cc6052b8d53c607d6876a2e5a3c44b607a324d4", overwrite=TRUE)
```

## Load data

**Task** Create a codeblock and load the WV slice of the PPP loan data, which is in the data folder and is called lab_11.rds.

```{r}
wv_ppp <- read_rds("data/lab_11.rds")
```

## Questions

```{r}
naics_two_digit <- wv_ppp %>%
  mutate(naics_code = str_sub(naics_code, start=1L, end=2L))
  
naics_21_only <- naics_two_digit %>%
  filter(naics_code == "21")

wv_loans_by_county <- naics_21_only %>%
  as_tibble() %>%
  group_by(project_county_name) %>%
  summarise(
    loans_by_county = sum(amount)
  ) %>%
  arrange(desc(project_county_name))

wv_county_population <- get_acs(geography = "county", 
              variables = c(population = "B01001_001"), 
              state = "WV",
              geometry = TRUE)

wv_county_population

naics_21_only

wv_county_population <- wv_county_population %>%
  rename(COUNTY = NAME) %>%
  mutate(COUNTY = toupper(str_remove_all(COUNTY,", West Virginia|County"))) %>%
  mutate(COUNTY = str_trim(COUNTY,side="both")) %>%
  rename(population = estimate) %>%
  select(COUNTY, population, geometry)

wv_county_population <- wv_county_population %>%
  rename(project_county_name = COUNTY)

wv_loans_21_per_100k <- wv_county_population %>%
  left_join(wv_loans_by_county) %>%
  mutate(loans_by_county = replace_na(loans_by_county,0)) %>%
  mutate(loans_per_100k = loans_by_county/population*100000) %>%
  arrange(desc(loans_per_100k))
```

```{r}
ggplot() + 
  geom_sf(data=wv_loans_21_per_100k, aes(fill=loans_per_100k)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") 
```

**Q1.** You are interested in doing some reporting on how West Virginia companies in the "Mining, Quarrying, and Oil and Gas Extraction" industry (https://www.naics.com/six-digit-naics/?code=21) used PPP loans during the pandemic (West Virginia is a big coal mining state).  

Then produce a choropleth map that shows West Virginia counties shaded according to the total amount of loans to that industry per 100,000 population.

Use the magma color palette, and shade on a "log" scale. Also produce a table that shows the total amount of loans per 100,000 population for each county, ranked from highest to lowest, along with the total amount of loans and the total number of loans.

Then answer these questions:

* Based on the map, what region of the state appears to have gotten more loans in this industry? What geographic pattern do you notice in the data?
* Based on the table, what county got the highest total loan amount? What was the total amount per 100K population, the total amount of loans (not population adjusted), how many loans were there, and what's the population?

```{r}
wv_loans_21_per_100k %>%
  arrange(desc(loans_by_county))
```

**A1.** * Based on the map, the north central region of West Virginia appears to have gotten the most loans in the industry. That region also has the county (Ritchie) with the most loans per industry per 100,000 in the state. The geographic pattern I notice in the data is that we can essentially break the state into four geographic regions - east, central, northwest and southwest. The eastern portion of the state was the section that had the lowest amount loans - there even were ome fields that we changed to zero in the data becasue they were N/A. The central region seems to be the most uniform one. Almost all of the loans in the central region fall within the orange block, which is the second highest in terms of loans per 100,000K. Then the northwest is the region with the most loans per 100k, while the southwest is the opposite of its northern neighbor and is mostly in the lower portion of the map in terms of loans per 100k.


Based on the table, the county that got the highest total loan amount was Harrison. The total amount per 100K population is 14,607,650, the total amount of loans is 9,919,763 and the population is 67,908  

**Q2.** You want to understand WHY the spatial pattern you saw above exists.  So you grab a shapefile with the latitude and longitude coordinates of every mine permit in West Virginia from the state Department of Environmental Protection.

Here's the url: https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json

(It's in a slightly different format than the .shp file in the lab. It's another type of shapefile called "geojson".  It will work just the same, though, despite that difference. You can read it directly into R without manually downloading it, by putting the url in the st_read() function we learned about in this week's prelab.)

This data includes both currently operating mines and old ones that have closed.  You want to examine only mines that have a permit expiration date after Sept. 1, 2021.

Make a map with a point for each one of these mine permits and put a blank map with the outline of West Virginia counties (which you can get from Tigris) behind it.

What spatial pattern do you see? How does it compare to the spatial pattern in the first map you made?

```{r}
library(lubridate)
```

```{r}
mining_permits <- st_read("https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json")

mining_permits <- mining_permits %>%
  mutate(expiredate = ymd(expiredate)) %>%
  filter(expiredate > "2021-09-01")
  
wv_counties <- counties %>%
  filter(STATEFP == "54")

ggplot() + 
  geom_sf(data=wv_counties) + 
  geom_sf(data=mining_permits) +
  theme_minimal()


```

**A2.**

**Q3.** To confirm the spatial comparison you identified in question 2, let's create a new map that layers the mine location points on top of our PPP choropleth map we created in question 1. What do you notice about the number of mines in the county that has the highest total loan amount per 100K?

**A3.**

**Q4.** Starting from the original WV ppp dataframe, examine the loans to companies in "Mining, Quarrying, and Oil and Gas Extraction" in the county with the highest total loan amount per 100K (you can use the table you created in question 1 to identify the top county).

```{r}
ritchie <- wv_ppp %>%
    mutate(naics_2 = str_sub(naics_code, start=1L, end=2L)) %>%
    filter(naics_2 == "21") %>%
    filter(project_county_name == "RITCHIE") %>%
    arrange(desc(amount))
```

**A4.**
* What single company accounts for the biggest chunk of the total loan amount?
    HALL DRILLING, LLC
* What specific industry are they in, based on their 6-digit NAICS code? 
    Drilling Oil and Gas Wells (213111)
*What services do they provide?
    -Trucking
    -Drilling
    -Product Installation
    -Construction
    -Roustabout Services
* How many jobs did they save, according to the data?
    387
* What's your hypothesis for why a company in their specific industry would need to seek a PPP loan during the pandemic?
    It's curious. There are several services that you would think would still be around during the pandemic, specifically           trucking. People were ordering a lot of goods during the pandemic, so I feel like trucking would still be a key need during     the pandemic to transport products. On the other hand, I have to assume that not many people were having people come in and     install equipment in their homes, so the product installation service would have been hit hard. This company also works with     a lot of expensive equipment, so they probably needed the loans to keep up with the maintenance costs of pricey equipment.

-30-
