---
title: "R Exercises, Part 3"
author: "Derek Willis"
date: "7/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More Summarization in R

We use the `tidyverse` library to load and calculate basic summaries of data in R. These exercises are focused on two of the basic summary verbs: `group_by` and `summarize` from the tidyverse, plus `filter` and `arrange`. Let's start by loading the tidyverse and some data using the `read_csv()` function.

```{r}
library(tidyverse)
time_stats <- read_rds("data/players.rds")
```

This data season statistics for NCAA women's college basketball players in 2021-22. The data has both raw totals for things like points, rebounds and assists along with some calculated percentages and ratios. The column names should be pretty self-explanatory. These questions will have you using filter, group_by, summarize and arrange (maybe not all at the same time, though). And we'll try making a scatterplot, too!

## Questions

1. Which Big Ten players had an assist-to-turnover ratio of at least 2 and played at least 500 minutes (`mp`)?
```{r}
big_10_ast_tov_500_mins <- time_stats %>%  
  filter(ncaa_conference == 'Big Ten', ast_tov_ratio >= 2.0, mp >= 500) %>%
  group_by(player, team)
```

#A: 	
Katie Benzan (Maryland), Gadiva Hubbard (Minnesota), Veronica Burton (Northwestern), Rikki Harris (OSU)

2. Expand the above code to include players in any conference with an assist-to-turnover ratio of at least 2 and played at least 500 minutes and make a scatterplot, using assists and turnovers as your X & Y axis (you made a scatterplot [here](https://github.com/iadecker/data_journalism_2022_spring/blob/main/labs/lab_09/lab_09.Rmd#L163))

```{r}
all_conference_ast_tov_500_mins <- time_stats %>%  
  filter(ast_tov_ratio >= 2, mp >= 500, ast >= 150) 

#for line 39, ast >= 150 is to help clean up the scatterplot, makes it less busy.

all_conference_ast_tov_500_mins %>%
  ggplot(aes(x=ast, y=tov, label=player)) + 
  geom_point() + 
  geom_text() +
 labs(
    title="Players with assist-to-turnover ratio of at least 2 and at least 500 minutes",
    x = "Assists",
    y = "Turnovers",
    caption = "Source: Derek Willis"
    ) 
```

3. Let's identify pure shooters: among players who made at least 80 three-point field goals, who had the highest 3-pointers made per minute played? (You'll need to calculate 3 pointers made per minute.)

```{r}
pure_shooters <- time_stats %>%
  mutate(made_threes_per_min = fg3m/mp) %>%
  filter(fg3m >= 80) %>%
  arrange(desc(made_threes_per_min))
```

4. Count how many players at each position played at least 200 minutes and never attempted a three-point shot?

```{r}
two_hundred_mp_w_no_threes <- time_stats %>%
  filter(mp >= 200, fg3a == 0) %>%
  group_by(position) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

#Just curious about who these three players are
na <- time_stats %>%
  filter(mp >= 200, fg3a == 0) %>%
  filter(is.na(position))
```

#A:
F: 186
C: 67
G: 11
N/A: 3

5. Who are the guards that did this? Are any from the Big Ten?

```{r}
guards <- two_hundred_mp_w_no_threes %>%
  filter(position == 'G') %>%
  group_by(ncaa_conference)

#OR, we could have done this above!

two_hundred_mp_w_no_threes <- time_stats %>%
  filter(mp >= 200, fg3a == 0, position == 'G') %>%
  group_by(player) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

#A: 
There were a total of 11 guards who fell under this category, with Sydney Hilliard of Wisconsin the only Big 10 player.

Here are the others:
Camille Gray (LIU)
Camryn Brown (UCLA)
Endia Holliday (Western Carolina)
Grace Efosa-Aguebor (Providence)
Jayla Ruffus-Milner (Pepperdine)
Julia Blackshell-Fair (Houston)
Liatu King (Pitt)
Mar'shaun Bostic (Auburn)
Mikayla Coombs (Georgia)
Tayra Eke (LIU)